---
title: "Global prediction of soil saturated hydraulic conductivity using random forest in a Covariate-based Geo Transfer Functions (CoGTF) framework"
authors: 
date: "Surya Gupta, Peter Lehmann, Sara Bonetti, Andreas Papritz and Dani Or"
output: 
  github_document:
    toc: true
---

Soil saturated hydraulic conductivity (Ksat) is one of the prominent soil hydraulic properties used in the modeling of land surface processes. Ksat is often derived using limited dataset and soil basic properties likely soil texture, bulk density) by means pedotransfer functions (PTFs). We propose here an integrated Predictive Soil Modeling (PSM) framework where soil variables are combined with RS-based covariates using the Random Forest method. We refer to this approach as the “Covariate-based Geo Transfer Functions” (CoGTF). Here, the objective of this report to show the methods used to develop the CoGTF with R code and stepwise description.   


SoilKsatDB [link](https://doi.org/10.5281/zenodo.3752721)

CoGTF global Ksat maps [link](https://doi.org/10.5281/zenodo.3934853)

To cite this maps please use:

Gupta, S., Hengl, T., Lehmann, P., Bonetti, S., Papritz, A. and Or, D.: [Global prediction of soil saturated hydraulic conductivity using random forest in a Covariate-based Geo Transfer Functions (CoGTF) framework](https://www.essoar.org/doi/10.1002/essoar.10503663.1). manuscript submitted to Journal of Advances in Modeling Earth Systems (JAMES).


```{r}
library(caret)
library(randomForest)
library(ranger)
library(mlr)
library(tibble)
library(raster)
library(sp)
library(rgdal)
library(hexbin)
library(lattice)
library(RColorBrewer)
library(viridis)
library(Metrics)

ksat_df<-read.csv("E:/Ksat_dataset_mapping.csv")

## Unique IDs for 5degrees by 5 degrees



source("E:/OpenLandMap/R/saveRDS_functions.R")
source("E:/OpenLandMap/R/LandGIS_functions.R")

## saveRDS_functions.R and LandGIS_functions.R available at https://github.com/Envirometrix/LandGISmaps/tree/477460d1d0099646c508f65e68769b9edf050ce8/functions

## 3D modeling (see Hengl, T., & MacMillan, R. A. (2019). Predictive soil mapping with R. Lulu. com.)

dfs <- hor2xyd(ksat_df, U="hzn_top", L="hzn_bot")


I.vars = make.names(unique(unlist(sapply(c("s.no","clm_", "dtm_", "lcv", "veg_", "olm_c", "olm_s", "olm_bd", "FID_Fish_", "DEPTH"), function(i){names(dfs)[grep(i, names(dfs))]}))))

t.vars = c("log_ksat")
sel.n <- c(t.vars,I.vars)
sel.r <- complete.cases(dfs[,sel.n])
PTF_temp2 <- dfs[sel.r,sel.n]



```

# Selection of random grids for spatial cross-validation

```{r}
## we selected 3 sets

##Set1
set.seed(16)
chosen <- sample(unique(PTF_temp2$FID_Fish_n), 40)

ff<-subset(PTF_temp2, FID_Fish_n %in% chosen)

final<-PTF_temp2[!(PTF_temp2$FID_Fish_n %in% ff$FID_Fish_n),]

set.seed(10)
chosen <- sample(unique(final$FID_Fish_n), 28)

ff1<-subset(final, FID_Fish_n %in% chosen)

final1<-final[!(final$FID_Fish_n %in% ff1$FID_Fish_n),]


set.seed(16)
chosen <- sample(unique(final1$FID_Fish_n), 16)

ff2<-subset(final1, FID_Fish_n %in% chosen)

final2<-final1[!(final1$FID_Fish_n %in% ff2$FID_Fish_n),]


set.seed(17)
chosen <- sample(unique(final2$FID_Fish_n), 53)

ff3<-subset(final2, FID_Fish_n %in% chosen)

final3<-final2[!(final2$FID_Fish_n %in% ff3$FID_Fish_n),]

##Set2

set.seed(34)
chosen <- sample(unique(PTF_temp2$FID_Fish_n), 58)

ff<-subset(PTF_temp2, FID_Fish_n %in% chosen)

final<-PTF_temp2[!(PTF_temp2$FID_Fish_n %in% ff$FID_Fish_n),]

set.seed(39)
chosen <- sample(unique(final$FID_Fish_n), 29)

ff1<-subset(final, FID_Fish_n %in% chosen)

final1<-final[!(final$FID_Fish_n %in% ff1$FID_Fish_n),]


set.seed(57)
chosen <- sample(unique(final1$FID_Fish_n), 27)

ff2<-subset(final1, FID_Fish_n %in% chosen)

final2<-final1[!(final1$FID_Fish_n %in% ff2$FID_Fish_n),]


set.seed(71)
chosen <- sample(unique(final2$FID_Fish_n), 12)

ff3<-subset(final2, FID_Fish_n %in% chosen)

final3<-final2[!(final2$FID_Fish_n %in% ff3$FID_Fish_n),]

##Set3

set.seed(79)
chosen <- sample(unique(PTF_temp2$FID_Fish_n), 29)

ff<-subset(PTF_temp2, FID_Fish_n %in% chosen)

final<-PTF_temp2[!(PTF_temp2$FID_Fish_n %in% ff$FID_Fish_n),]

set.seed(85)
chosen <- sample(unique(final$FID_Fish_n), 36)

ff1<-subset(final, FID_Fish_n %in% chosen)

final1<-final[!(final$FID_Fish_n %in% ff1$FID_Fish_n),]


set.seed(100)
chosen <- sample(unique(final1$FID_Fish_n), 31)

ff2<-subset(final1, FID_Fish_n %in% chosen)

final2<-final1[!(final1$FID_Fish_n %in% ff2$FID_Fish_n),]


set.seed(117)
chosen <- sample(unique(final2$FID_Fish_n), 27)

ff3<-subset(final2, FID_Fish_n %in% chosen)

final3<-final2[!(final2$FID_Fish_n %in% ff3$FID_Fish_n),]



df1<-ff
df2<-ff1
df3<-ff2
df4<-ff3
df5<-final3



Train1<- rbind(ff, ff1, ff2, ff3)

Train2<- rbind (ff1, ff2, ff3,final3)

Train3<- rbind(ff2, ff3,final3, ff)

Train4<- rbind(ff3,final3, ff,ff1)

Train5<- rbind(final3, ff,ff1, ff2)

```
# Selection of Covaraites

```{r}
grid <- list.files("E:/maps_tests/new_layers/layers_RS/" , pattern = "*.tif$")
All_cov <- raster::stack(paste0("E:/maps_tests/new_layers/layers_RS/", grid))

set.seed(2) 
fm.ksat <- as.formula(paste("log_ksat~ ",paste(names(All_cov), collapse = "+")))
fm.ksat

```
# Model fitting
```{r}
set.seed(2) 
rm.ksat <- Train1[complete.cases(Train1[,all.vars(fm.ksat)]),]
m.ksat <- ranger(fm.ksat, rm.ksat, num.trees=200, mtry=6, quantreg = TRUE)
m.ksat

df5$prediction<- predict(m.ksat,df5)$predictions

RMSE(df5$prediction, df5$log_ksat)

## Ist_part is computed

rm.ksat1 <- Train2[complete.cases(Train2[,all.vars(fm.ksat)]),]
m.ksat1 <- ranger(fm.ksat, rm.ksat1, num.trees=200, mtry=6, quantreg = TRUE)
m.ksat1

df1$prediction<- predict(m.ksat1,df1)$predictions

RMSE(df1$prediction, df1$log_ksat)

## 2nd_part is computed
rm.ksat2 <- Train3[complete.cases(Train3[,all.vars(fm.ksat)]),]
m.ksat2 <- ranger(fm.ksat, rm.ksat2, num.trees=200, mtry=6, quantreg = TRUE)
m.ksat2

df2$prediction<- predict(m.ksat2,df2)$predictions

RMSE(df2$prediction, df2$log_ksat)

## 3rd_part is computed

rm.ksat3 <- Train4[complete.cases(Train4[,all.vars(fm.ksat)]),]
m.ksat3 <- ranger(fm.ksat, rm.ksat3, num.trees=200, mtry=6, quantreg = TRUE)
m.ksat3
df3$prediction<- predict(m.ksat3,df3)$predictions

RMSE(df3$prediction, df3$log_ksat)

## 4th_part is computed
rm.ksat4 <- Train5[complete.cases(Train5[,all.vars(fm.ksat)]),]
m.ksat4 <- ranger(fm.ksat, rm.ksat4, num.trees=200, mtry=6, quantreg = TRUE)
m.ksat4

df4$prediction<- predict(m.ksat4,df4)$predictions

RMSE(df4$prediction, df4$log_ksat)

Final_data<- rbind(df1,df2,df3,df4,df5)

dd<- aggregate(Final_data[, 1:33], list(Final_data$s_no), mean)




RMSE(dd$prediction,dd$log_ksat)



ccc = DescTools::CCC(dd$prediction,dd$log_ksat, ci = "z-transform", conf.level = 0.95, na.rm=TRUE)$rho.c
ccc


dd$log_ksat1<- 10^dd$log_ksat
dd$prediction1<- 10^dd$prediction

hexbinplot(log_ksat1~ prediction1, 
           panel = function(x, y, ...){
             panel.hexbinplot(x, y, ...)
             panel.loess(x, y,span = 2/3, col.line = "blue",type="l", lty=2, lwd = 4)
             panel.abline(c(0, 1),lwd = 2)
           },
           data = dd,xlab = "Predicted Ksat [cm/day]", ylab = "Measured Ksat [cm/day]",cex.axis = 4, aspect="1", xbins=30, colramp = function(n) {viridis (8,  alpha = 1, begin = 0, end = 1, direction = -1,option = "C")},xlim=c(0.1,10000), ylim=c(0.1,10000),
           scales=list(
             x = list(log = 10, equispaced.log = FALSE), 
             y = list(log = 10, equispaced.log = FALSE)
           ),
           font.lab= 6, cex.labels = 1.2,font.axis = 2,colorcut=c(0,0.01,0.03,0.07,0.15,0.25,0.5,0.75,1) )


##Fitting final model

#rm.ksat <- PTF_temp2[complete.cases(PTF_temp2[,all.vars(fm.ksat)]),]
#m.ksat <- ranger(fm.ksat, rm.ksat, num.trees=200, mtry=6, quantreg = TRUE)
#m.ksat


##Importance variable
#m.ksat <- randomForest(fm.ksat, rm.ksat, num.trees=200, mtry=6, quantreg = TRUE)

#varImpPlot(m.ksat, sort=TRUE, n.var=min(30, nrow(m.ksat$importance)))

## Then Produced the final CoGTF map

#p2 = predict(All_cov,m.ksat, progress='window',type = "response",fun = function(model, ...) predict(model, ...)$predictions)

#writeRaster(p2, "/home/step/data/OpenLandMap/Final_selected_covariates/New_raster_layer/Final_RF_0cm.tif")

```

